---
title: "Simulation using the exponential distribution"
author: "Vaibhav Palve"
output: "pdf_document"
---

```{r}
library(ggplot2)
```

Now, set the variables as defined in the problem. number of values (n) = 40       lambda = 0.2 number of iterations, as least 1000. nosim=1000.
Initialize the variables.
```{r}
nosim <- 1000;
n <- 40;
lambda <- 0.2;
```

Using the rexp function, along with the matrix function develop a dataset with the mean and lambda specified above

Also, the theoretical mean is given as 1 / lambda or 1 / 0.2.
The actual mean of the generated data can be calculated by using the apply &      mean functions to obtain a mean for each row and then taking the mean of those    numbers.

```{r}
 ## values
simdata <- matrix(rexp(nosim * n, rate=lambda), nosim);
simdata_means <- apply(simdata, 1, mean);
hist(simdata_means);

theoretical_mean <- 1/lambda;

print (paste("Theoretical center of the distribution = ",theoretical_mean));
print (paste("Actual center of the distribution based on the simulations = ", 
             round(mean(simdata_means), 3)));

theoretical_var <- (1/lambda)^2/n;
theoretical_sd <- 1/lambda/sqrt(n);
print (paste("Theoretical variance = ", theoretical_var));
print (paste("Actual variance based on the simulations=",
             round(var(simdata_means), 3)));

print (paste("Theoretical standard deviation = ", 
                     round(theoretical_sd, 3)));

print (paste("Actual standard deviation based on the simulations = ", 
                     round(sd(simdata_means), 3)));

plotdata <- data.frame(simdata_means);
m <- ggplot(plotdata, aes(x = simdata_means)); 
m <- m + geom_histogram(aes(y=..density..), colour="black", 
                                fill = "blue")
m + geom_density(colour="red", size=1);
```

---
To answer question 1 - I have demonstrated the actual mean compared to the theroetical mean. You can check the value in the output above.
To Answer question 2 - I have demonstrated the actual standard deviation and variance compared to their theoretical values. 
To answer question 3 - I graphed and was able to show that both the calculated curve is very close to the theoretical normal curve therefore proving the Central Limit Theory.

